---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1. ggplot
1. Consider the following code. Using your own words, explain why the points are not green? (1 point)
```{r}
require(ggplot2)
ggplot(data = airquality) + 
  geom_point(mapping = aes(x = Wind, y = Ozone, color = "green"))
```

> Because the code color = "green" should not be in the same bracket with x and y. It should be geom_point(mapping = aes(x = Wind, y = Ozone), color = "green").

```{r}
require(ggplot2)
ggplot(data = airquality) + 
  geom_point(mapping = aes(x = Wind, y = Ozone), color = "green")
```

2. Map a continuous variable to color, size, and shape. In words, explain why these aesthetics behave differently for categorical vs. continuous variables. (1 point)
```{r}
# categorical variables
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, color = class))

ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, size = class))

ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, shape = class))

# discrete variables
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, color = cty))

ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, size = cty))

#ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, shape = cty))
```

> I mapped categorical variable class and continuous variable cty to to color, size, and shape respectively. Categorical variables contain a finite number of categories or distinct groups whereas Continuous variables are numeric variables that have an infinite number of values between any two values. In terms of mapping to color, when we used continuous variable as grouping factor, the color bar is gradient color from light to dark instead of totally different color for categorical variable. In terms of mapping to shape, the sizes of the points vary continuously as a function of their size. In terms of mapping to shape, it gives an error for continuous value because shape doesn't have an order. 

3. In words, explain what happens if you map the same variable to multiple aesthetics? (1 point)

> It is pretty redundant. The plot will provide the same information in many different ways but they are repetitive. I think the most important factor for a good plot and figure is effectively and efficiently to delivery the message.

4. What happens if you map an aesthetic to something other than a variable name, like aes(colour = Temp < 70)? Note, you’ll also need to specify x and y (2 point)
```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, colour = cty < 20))
```

>I mapped cty<20. The result of cty<20 is a logic variable. If cty<20, it will take value of TRUE. If ctyis greater or equal to 20, it will take value of FALSE. In the plot, TRUE and FALSE will have different color. This is a good way to seperate data based on some critical values.

## 2. Simulation study
1. Generate a single simple random sample of length n <- 50 from the binomial distribution with the parameters k <- 10, p <- 0.4 (1 point).
```{r}
n <- 50
k <- 10
p <- 0.4
sample <- rbinom(n,size = k, prob = p)
```

2. Write a function that takes a sample as its input and returns a vector containing the estimates of k and p given above (3 points).
```{r}
estimate <- function(sam){
  mean_sam <- mean(sam)
  var_sam <- var(sam)
  n <- length(sam)
  k_hat <- mean_sam^2/(mean_sam-(n-1)/n*var_sam)
  p_hat <- mean_sam/k_hat
  return(c(k_hat = k_hat, p_hat = p_hat))
}
estimate(sample)
```

3. Generate N <- 1000 samples of size n <- 50 (as in the first question) and calculate N <- 1000 estimates of k and N <- 1000 estimates of p (4 points if no loops are used in the code, 2 points if any loop (for, while, repeat, etc.) is used in the code ).
```{r}
N <- 1000
sampleN1 <- matrix(rbinom(N*n,k,p),n,N)
estimateN1 <- t(apply(sampleN1,2, FUN=estimate))
head(estimateN1)
```

4. Repeat Question 3 when n <- 100 and when n <- 250 (2 points).
```{r}
n <- 100
sampleN2 <- matrix(rbinom(N*n,k,p),n,N)
estimateN2 <- t(apply(sampleN2,2,FUN=estimate))
head(estimateN2) #just show the first 6 rows

n <- 250
sampleN3 <- matrix(rbinom(N*n,k,p),n,N)
estimateN3 <- t(apply(sampleN3,2, FUN=estimate))
head(estimateN3) #just show the first 6 rows
```

5. Estimate the bias and the mean squared error (MSE) of $\hat k$ and the bias and the MSE of $\hat p$ for each sample size (`n <- 50`, `n <- 100` and `n <- 250`). Do the estimators seem to overestimate or underestimate the parameters? How do the bias and the MSE change when the sample size increases? .
```{r}
kbiasmse <- function(x){
  k_bias <- mean(x)-k
  k_mse <- var(x) + k_bias^2
  return(c(k_bias=k_bias,k_mse=k_mse))
}

pbiasmse <- function(x){
  p_bias <- mean(x)-p
  p_mse <- var(x) + p_bias^2
  return(c(p_bias=p_bias,p_mse=p_mse))
}


k_data <- cbind(estimateN1[,1],estimateN2[,1],estimateN3[,1])
kmatrix <- t(apply(k_data, 2, kbiasmse))
rownames(kmatrix) <- c("n=50","n=100","n=250")
print(kmatrix)

p_data <- cbind(estimateN1[,2],estimateN2[,2],estimateN3[,2])
pmatrix <- t(apply(p_data, 2, pbiasmse))
rownames(pmatrix) <- c("n=50","n=100","n=250")
print(pmatrix)
```
The estimators seem to overestimate the parameters because the biases are always positive.
We can see that with n getting larger, the bias and mse of the estimation of k and p tend to be closer to 0 except the bias of p under n=250.

6. a. Make a single plot using `ggplot2` that contains three box plots of the estimates of the parameter `k` when `n <- 50`, `n <- 100`, `n <- 250` (the first from the left box plot has to describe the estimates when `n <- 50`, the second from the left box plot has to describe the estimates when `n <- 100` and the third from the left box plot has to describe the estimates `n <- 250`). Include the true value of the parameter as a horizontal line (`geom_hline()` and use the argument `color`) and label the plot appropriately.
```{r}
library(ggplot2)
kvalue <- matrix(c(estimateN1[,1],estimateN2[,1],estimateN3[,1]),N*3,byrow = FALSE)
label <- factor(c(rep("n=50",N),rep("n=100",N),rep("n=250",N)), levels <- c("n=50", "n=100","n=250"))
kdataframe <- data.frame(kvalue, label)
ggplot(data = kdataframe, mapping = aes(x = label, y = kvalue)) +
  geom_boxplot()+
  geom_hline(yintercept = k,color= "red") +
  labs(title="Box plots of estimators k per n",x="Sample size", y = "Estimators k")
```

  b. $\hat k$ can obtain values that are far away from the true value of the parameter when the sample size is small and the box plots might not be particularly informative in such a situation. Remove the estimates from the plot that are outside of the interval $[0,50]$ so that the box plots are more informative.
```{r}
kdataframe_clean <- kdataframe[kdataframe$kvalue >= 0 &kdataframe$kvalue <= 50,]
ggplot(data = kdataframe_clean, mapping = aes(x = label, y = kvalue)) +
  geom_boxplot()+
  geom_hline(yintercept = k,color= "red") +
  labs(title="Box plots of estimators k per n",x="Sample size", y = "Estimators k")
```

  c. Make the same plot with three box plots for the estimates of the parameter `p` (b part does not apply here).
```{r}
pvalue <- matrix(c(estimateN1[,2],estimateN2[,2],estimateN3[,2]),N*3,byrow = FALSE)
label <- factor(c(rep("n=50",N),rep("n=100",N),rep("n=250",N)), levels <- c("n=50", "n=100","n=250"))
pdataframe <- data.frame(pvalue, label)
ggplot(data = pdataframe, mapping = aes(x = label, y = pvalue)) +
  geom_boxplot()+
  geom_hline(yintercept = p,color= "red") +
  labs(title="Box plots of estimators k per n",x="Sample size", y = "Estimators k")
```

  d. Describe how both of these plots change when the sample size increases.
  
>When the sample size increases, the outliers are fewer, and the range of boxplot tends to to smaller, and the IQR of boxplot tends to be smaller, and the median is closer to the true value.

## Extra credit (3 points)
Reproduce the plots in 2) using base R plotting functions.
```{r}
kvalue <- matrix(c(estimateN1[,1],estimateN2[,1],estimateN3[,1]),N*3,byrow = FALSE)
label <- c(rep(50,N),rep(100,N),rep(250,N))
dataframe <- data.frame(kvalue, label)

boxplot(kvalue~label, data=dataframe,
        main = "Box plots of estimators k per n",
        xlab = "Sample size",
        ylab = "Estimators k")
abline (h= k,col= "red")

dataframe_clean <- dataframe[kdataframe$kvalue >= 0 &kdataframe$kvalue <= 50,]
boxplot(kvalue~label, data=dataframe_clean,
        main = "Box plots of estimators k per n",
        xlab = "Sample size",
        ylab = "Estimators k")
abline (h= k,col= "red")
```

